Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-19 doc/manuscript/translocation_bioarxiv.tex
Adjust manuscript formatting

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

@@

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. One possibility is rapid
evolution that restores positive growth to declining populations
(``evolutionary rescue''). Individuals from these rescued populations
could also be used to reestablish exinct populations. We evaluated both
possibilities in a once-common frog species driven to the brink of
extinction by a disease that has devastated global amphibian
biodiversity and for which few options exist to reverse disease impacts.
We found that frogs have evolved increased resistance to infection,
facilitating population recovery. In addition, introduction of frogs
from these populations allowed reestablishment despite the ongoing
presence of disease. Such reintroductions could play an important role
in biodiversity conservation in our rapidly changing world.}

@@

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

\dropcap{H}uman activities are intensifying threats to global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. To persist, affected
taxa will often need to adapt to their altered environment or face
decline and extinction. ``Evolutionary rescue'' describes rapid
evolutionary change that increases the frequency of adaptive alleles and
thereby restores positive population growth \citep{carlson2014}. In
addition to preventing extinction of declining populations, this process
could also result in the seeding of nearby empty habitats and allow
population reestablishment and recovery. Although this overall scenario
is potentially of critical importance for the conservation of an
ever-growing number of threatened taxa, descriptions of its operation in
the wild are lacking.

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa
\citep{hewson2014, samuel2015, scheele2019, cunningham2021}. Amid this
global loss of biodiverity, some populations decimated by disease have
subsequently recovered to varying extents
\citep{newell2013, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance (see
\citep{medzhitov2012} for definitions). However, linking these
evolutionary responses to population recoveries is difficult and few
clear examples exist. In addition to promoting recovery of
disease-impacted populations, the resistant/tolerant individuals in such
populations could potentially be used to reestablish populations that
were extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

In the current study, we combine results from genomic analyses, a
15-year landscape-scale frog reintroduction effort, and a detailed
population model to provide an unprecedented example of how evolution in
response to disease allows the successful reestablishment of extirpated
MYL frog populations. Specifically, we used exome capture methods to
show that MYL frogs have undergone substantial evolutionary change
associated with Bd exposure and that might provide a genetic basis for
the increased resistance/tolerance to Bd infection observed in naturally
recovering populations \citep{knapp2016}. We then use the reintroduction
of resistant/tolerant frogs to 24 sites and results from long-term
capture-mark-recapture (CMR) surveys to describe the dynamics and
drivers of population establishment. These data are also used to
parameterize a population model that extends our inferences of
population establishment well beyond the temporal extent of the CMR
data. Collectively, these results provide one of the first examples of
how evolutionary rescue and the reintroduction of resistant/tolerant
animals can allow the large-scale recovery of disease-impacted species.

@@

To determine whether MYL frog populations show an evolutionary response
to Bd, we compared frog genomes between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in four populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in five populations that experienced
a Bd epizootic during the past several decades and have since recovered
to varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. These recovering populations exist in
an enzootic state \citep{briggs2010}, characterized by high Bd
prevalence and, in adults, moderate Bd infection intensities (``load'')
that are typically well below the level expected to cause mortality
\citep{vredenburg2010}.

@@

Individual frogs clustered into 3 separate groups in principal component
space (Figure S1A), and clusters reflected the
strong signature of isolation by distance that is characteristic of MYL
frogs \citep{rothstein2020, poorten2017}. Importantly, each cluster
contained at least one population from both the naive and recovering
groups, allowing us to distinguish allelic associations of individuals
sampled in the 2 population types versus allelic associations resulting
from population structure and genetic drift.

Results from the individual variant and splined window analyses show
that recovering populations have signatures of strong selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S1B,C). One of the 7 identified genes did not have an associated
annotation. Frequency differences between the naive and recovering
populations ranged from 41\% to 86\%. Most of these SNPs showed
frequency differences in only a subset of the sampled populations
(Figure~\ref{fig-allelefrequencies} A, B), but the SNP in the RIN3 gene
showed consistent differences in frequencies across all populations
(Figure~\ref{fig-allelefrequencies} C). This is suggestive of strong -
and parallel - selection at this locus associated with exposure of frogs
to Bd across multiple populations.

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan} A, B). Of these, 9 regions were
outliers for both metrics and 2 of these shared regions also contained
outlier SNPs. A total of 33 annotated genes were found in the 9 shared
regions (SI). The largest \(\pi_{diff}\), indicative of directional
selection, occurred in a 163kb region on Contig19, 12.9Mb upstream of
the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan} C). This region
contains approximately 500 SNPs and one annotated gene called
``interferon-induced very large GTPase 1-like'' (GVINP1). Additionally,
a shared outlier region on Contig1 contained two complement factor genes
(C6 and C7). Interestingly, this region had a low \(\pi_{diff}\) value,
consistent with balancing selection. Finally, one shared outlier region
on Contig8 contained one outlier SNP (TCF19) and was within 360kb of
another outlier SNP (VARS) (Figure~\ref{fig-spline-manhattan} D,
Figure S2). This region (854kb from the beginning of
the outlier window to the VARS SNP) contained a total of 8 annotated
genes. In \emph{Xenopus}, five of these genes occur in the extended
major histocompatibility complex (MHC) Class I region (FLOT1, TUBB,
MDC1, CCHCR1, TCF19) and three occur in the extended MHC Class III
region (HSP70, LSM2, VARS) \citep{ohta2006}. Therefore, this region
under selection is part of the extended MHC Class I and III complex and
shows synteny with other amphibian genomes.

@@

To determine whether \emph{R. sierrae} from recovering populations could
be used to reestablish extirpated populations, we conducted 24
reintroductions in Yosemite National Park (2006--2020). Each of the
reintroductions involved collection of adult frogs from 1 of 3
recovering, Bd-positive ``donor'' populations (Figure S3) and translocating them to 1 of 12 nearby recipient sites
\citep{seddon2014}. The donor populations included 2 of the 5 recovering
populations in the frog evolution study described above
(Figure~\ref{fig-allelefrequencies} A: population 1 and 4). Following
translocation, adult survival and recruitment of new adults were
estimated from CMR surveys and counts of tadpoles and juveniles were
obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1 year following translocation
indicate that survival was highly variable between recipient sites, but
relatively constant within recipient sites (for the subset of sites that
received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability over time to choose recipient sites
with higher habitat quality for \emph{R. sierrae}. Finally, Bd loads
were fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5,
see SI text for additional details) \citep{joseph2018, vredenburg2010}.

@@

Even when incorporating demographic stochasticity and environmental
stochasticity in the probability of juvenile recruitment to the adult
class \(\omega\) (the transition that we expect to be the most subject
to environmental variability in the presence of Bd), populations with
high adult survival are likely to persist over a 50 year time horizon.
Our model predicted that, following a single introduction of 40 adult
individuals into a population and using the baseline value for
\(\omega\) of 0.3, the six populations with the highest adult survival
probabilities had 50-year extinction probabilities of 0 to 0.32;
Figure~\ref{fig-viability} B). This indicates strong potential for
long-term persistence in the presence of Bd and environmental
variability in recruitment (Figure~\ref{fig-viability} B). When we
decreased \(\omega\) from 0.3 to a conservative estimate of 0.15,
50-year extinction probability remained less than 0.5 for three
populations and was 0.79 to 0.94 for three additional populations
(Figure~\ref{fig-viability} B). In contrast, for the six populations
where yearly adult survival probability \(\sigma_{A_R} < 0.5\),
extinction probability over 50 years was always predicted to be 1. To
test the validity of our model predictions, we demonstrated that our
stochastic model could describe the general recovery trajectory of our
translocated population with the longest survey history
(Figure~\ref{fig-viability} C; population 70550, surveyed for 16 years).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but strategies to recover affected species are
generally lacking. Here, we used populations of resistant/tolerant
individuals to reestablish extirpated populations, despite ongoing
disease. Although previously proposed \citep{mendelson2019}, this
strategy has rarely been successfully implemented (see
\citep{joseph2018} for a small-scale example). In particular, although
numerous studies have documented signatures of selection in hosts
following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (i) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease and that some of
the genomic regions under selection are associated with genes related to
disease resistance/tolerance, and (ii) reintroduction of individuals
from these resistant/tolerant populations allows the successful
reestablishment of viable frog populations despite ongoing Bd infection.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also Figure S5 - before
translocation]{briggs2010, knapp2011}. The differences in Bd load of
frogs from naive and recovering populations are also observed in
controlled laboratory studies \citep[Figure 4]{knapp2016}. The observed
reduced susceptibility of frogs from recovering populations could in
theory be due to several factors including natural selection for more
resistant/tolerant genotypes and/or acquired immunity, but until now
evidence to evaluate the role of evolution was lacking.

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(\pi_{diff}\)) contains the gene GVINP1. Several previous studies have
found this gene to be differentially expressed during Bd infection
\citep[e.g.,][]{grogan2018, ellison2014} and in populations differing in
Bd susceptibility \citep{grogan2018}. This gene is also strongly linked
to disease in salmon, explaining a notable 20\% of the resistance
phenotype \citep{robledo2020, robledo2018}. We also identified a region,
characterized by high \emph{F\textsubscript{ST}} and low \(\pi_{diff}\),
that contained the complement genes C6 and C7. This could indicate that
balancing selection is acting in this region of the genome to favor a
diverse set of alleles, as is known for C6 in humans
\citep{soejima2005}. Finally, the RIN3 gene showed a consistent pattern
of allele frequency differences across all nine of the frog populations
sampled in this study, indicating consistent selection in populations
distributed across a wide geographic area. This gene is associated with
inflammation response and in \emph{Xenopus} is highly expressed in wound
stumps following amputation of tadpole tails \citep{fukazawa2009}. In
the SI, we provide additional details on these and other genes
identified in our study as under selection.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering
populations, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations, Bd
infection and load had detectable effects on the survival of adults and
may have influenced population establishment (sites referred to as
``Alpine'' and ``Subalpine'' in \citep{joseph2018} are identified as
``70550'' and ``70505'' in the current study). Applying similar analyses
to all 12 of our translocated populations would likely provide a broader
perspective of the ongoing effect of Bd. In addition to these important
but relatively subtle effects on adults, MYL frogs immediately following
metamorphosis (``metamorphs'') are highly susceptible to Bd infection
\citep{ellison2018} and experience high mortality. This high
susceptibility of metamorphs is documented for numerous species of
anurans, and may result from the poorly developed immune system
characteristic of this early life stage \citep{humphries2022}. In
naturally recovering and translocated MYL frog populations, we suggest
that the high mortality of metamorphs is an important limitation on
subsequent recruitment of new adults. Therefore, although adult MYL
frogs appear relatively resistant/tolerant, Bd infection continues to
have important limiting effects on recovering populations \citep[see
also][]{hollanders2022}.

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
\citep{scheele2019}. This unprecedented impact on global amphibian
biodiversity is compounded by the lack of any effective and broadly
applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of also
using animals from these recovering populations to reestablish
extirpated populations. As with MYL frogs, the feasibility and long-term
success of such efforts will depend on the availability of robust donor
populations containing individuals that have acquired via evolution the
adaptive alleles necessary to allow frog survival and population growth
in the presence of Bd. Despite the hopeful example of successful
reestablishment of MYL frogs via evolutionary rescue and
reintroductions, the challenge of recovering hundreds of Bd-impacted
amphibian species worldwide is a daunting prospect. Although we now have
a proven strategy to reestablish extirpated MYL frog populations,
recovery across their large historical range will require substantial
resources over many decades. The results of this study provide a hopeful
starting point for that endeavor, and other future efforts worldwide.

Evolutionary rescue may be increasingly critical for the survival of
wildlife populations in a rapidly changing world. The importance of
evolutionary rescue is known for a wide variety of taxa
\citep[e.g.,][]{carlson2014, mills2018, oziolor2019}, including in the
specific context of disease-impacted systems \citep{searle2020}.
However, obtaining evidence in wild animal populations for all the
causal links in this process (i.e., selection for adaptive alleles
\(\rightarrow\) increase in host fitness in the presence of the stressor
\(\rightarrow\) positive population growth and recovery) is difficult in
most systems. Nonetheless, we provide a compelling example from the wild
not only that evolutionary rescue can produce individuals that harbor
adaptive alleles and allow population recovery, but importantly that
these individuals can be used in translocations to reestablish
extirpated populations. We expect that such ``assisted evolutionary
rescue'' will be an essential tool in wildlife conservation in an era of
accelerating global change.

@@

To identify important predictors of frog survival following
translocation, we developed linear and multilevel models in a Bayesian
framework. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}, and used default, weakly informative
priors. Starting models included all relevant population-level effects
and, when data contained possible hierarchical structure, we included
group-level effects in the full model. Within each analysis, we
evaluated model fit using visualizations of leave-one-out (``LOO'')
probability integral transformations \citep{gelman2013, gabry2019}, and
compared fits of models with and without group-level effects, and with
differing group-level effects using LOO cross-validation
\citep{vehtari2016}. Both approaches utilized the loo package
\citep{vehtari2022}.

>>>>

\templatetype{pnasmathematics} % Choose template 
% {pnasmathematics} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

@@

% Please add a significance statement to explain the relevance of your work
\significancestatement{Understanding how species persist despite accelerating global change is
critical for the conservation of biodiversity. One possibility is rapid
evolution that restores positive growth to declining populations
(``evolutionary rescue''). Individuals from these rescued populations
could also be used to reestablish extinct populations. We evaluated both
possibilities in a once-common frog species driven to the brink of
extinction by a disease that has devastated global amphibian
biodiversity and for which few options exist to reverse disease impacts.
We found that frogs have evolved increased resistance to infection,
facilitating population recovery. In addition, introduction of frogs
from these populations allowed reestablishment despite the ongoing
presence of disease. Such reintroductions could play an important role
in biodiversity conservation in our rapidly changing world.}

@@

\dates{This manuscript was compiled on \today}
% \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

\dropcap{H}uman activities are intensifying threats to global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. To persist, affected
taxa will often need to adapt to their altered environment or face
decline and possible extinction. The process whereby rapid evolutionary
change increases the frequency of adaptive alleles and restores positive
population growth is known as ``evolutionary rescue''
\citep{carlson2014}. In addition to preventing extinction of declining
populations, individuals from rescued populations could also disperse,
or be reintroduced, to empty habitats and promote larger-scale recovery.
Although this overall scenario is potentially of critical importance for
the conservation of an ever-growing number of threatened taxa,
descriptions of its occurrence and effectiveness in the wild are
lacking.

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa, including echinoderms, mammals, birds, and
amphibians \citep{hewson2014, samuel2015, scheele2019, cunningham2021}.
Amid this global loss of biodiversity, some populations decimated by
disease have subsequently recovered to varying extents
\citep{newell2013, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance (see
\citep{medzhitov2012} for definitions). However, linking these
evolutionary responses to population recoveries is difficult and few
clear examples exist. In addition to promoting recovery of
disease-impacted populations, the resistant/tolerant individuals in such
populations could potentially be used to reestablish populations that
were extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

In the current study, we combine results from genomic analyses, a
15-year landscape-scale frog reintroduction effort, and a detailed
population model to provide an unprecedented example of how evolution in
response to disease allows the successful reestablishment of extirpated
MYL frog populations. Specifically, we used exome capture methods to
show that MYL frogs have undergone substantial evolutionary change
associated with Bd exposure, changes that might provide a genetic basis
for the increased resistance/tolerance to Bd infection observed in
naturally recovering populations \citep{knapp2016}. We then use the
reintroduction of resistant/tolerant frogs to 24 sites and results from
long-term capture-mark-recapture (CMR) surveys to describe the dynamics
and drivers of population establishment. These data are also used to
parameterize a population model that extends our inferences of
population establishment well beyond the temporal extent of the CMR
data. Collectively, these results provide one of the first examples of
how evolutionary rescue and the reintroduction of resistant/tolerant
animals can allow the large-scale recovery of disease-impacted species.

@@

To determine whether MYL frog populations show an evolutionary response
to Bd, we compared frog genomes between populations with contrasting
histories of Bd exposure. Specifically, we compared frog genomes sampled
in four populations that have not yet experienced a Bd-caused epizootic
(``naive'') \citep{zhou2015} versus in five populations that experienced
a Bd epizootic during the past several decades and have since recovered
to varying degrees (``recovering''; Figure~\ref{fig-allelefrequencies})
\citep{knapp2016, vredenburg2010}. These recovering populations exist in
an enzootic state \citep{briggs2010}, characterized by high Bd
prevalence and, in adults, moderate Bd infection intensity (``load'')
that is typically well below the level expected to cause mortality
\citep{vredenburg2010}. (See SI Appendix for additional details
regarding the study design.)

@@

Individual frogs clustered into 3 separate groups in principal component
space (Figure S1A), and clusters reflected the
species split (i.e., \emph{R. muscosa} versus \emph{R. sierrae}) and the
strong signature of isolation by distance that is characteristic of MYL
frogs \citep{rothstein2020, poorten2017}. Importantly, each cluster
contained at least one population from both the naive and recovering
groups, allowing us to distinguish allelic associations of individuals
sampled in the 2 population types versus allelic associations resulting
from population structure and genetic drift.

Results from the individual variant and splined window analyses show
that recovering populations have signatures of strong selection on
multiple regions of the genome. The analysis of individual variants
identified 11 ``outlier'' SNPs (i.e., showing significantly different
allele frequencies between naive versus recovering populations) from 7
distinct genes across 4 contigs (Figure S1B,C). One of the 7 identified genes (LOC108802036) did not have an
associated annotation. For the outlier SNPs, frequency differences
between the naive and recovering populations ranged from 41\% to 86\%.
Most of these SNPs showed frequency differences in only a subset of the
sampled populations (Figure~\ref{fig-allelefrequencies} A, B), but the
SNP in the RIN3 gene showed consistent differences in frequencies across
all populations (Figure~\ref{fig-allelefrequencies} C). This is
suggestive of strong - and parallel - selection at this locus associated
with exposure of frogs to Bd across multiple populations. (See SI
Appendix for additional description of these results.)

The splined window analysis identified 33 outlier regions for
\(\pi_{diff}\) and 58 outlier regions for \emph{F\textsubscript{ST}}
(Figure~\ref{fig-spline-manhattan} A, B). Of these, 9 regions were
outliers for both metrics (``shared regions'') and 2 of these shared
regions also contained one or more outlier SNPs (described above). A
total of 33 annotated genes were found in the 9 shared regions (SI
Appendix: Datasets S3, S4). The largest \(\pi_{diff}\), indicative of
directional selection, occurred in a 163kb region on Contig19, 12.9Mb
upstream of the RIN3 outlier SNP (Figure~\ref{fig-spline-manhattan} C).
This region contains approximately 500 SNPs and one annotated gene
called ``interferon-induced very large GTPase 1-like'' (GVINP1).
Additionally, a shared outlier region on Contig1 contained two
complement factor genes (C6 and C7). Interestingly, this region had a
large negative \(\pi_{diff}\) value, consistent with balancing
selection. Finally, one shared outlier region on Contig8 contained one
outlier SNP (TCF19) and was within 360kb of another outlier SNP (VARS)
(Figure~\ref{fig-spline-manhattan} D, Figure S2).
This region (854kb from the beginning of the outlier window to the VARS
SNP) contained a total of 8 annotated genes. In \emph{Xenopus}, five of
these genes occur in the extended major histocompatibility complex (MHC)
Class I region (FLOT1, TUBB, MDC1, CCHCR1, TCF19) and three occur in the
extended MHC Class III region (HSP70, LSM2, VARS) \citep{ohta2006}.
Therefore, this region under selection is part of the extended MHC Class
I and III complex and shows synteny with other amphibian genomes.

@@

To determine whether frogs from recovering populations could be used to
reestablish extirpated populations, we conducted 24 reintroductions in
Yosemite National Park (2006--2020). Each of the reintroductions
involved collection of adult frogs from 1 of 3 recovering, Bd-positive
``donor'' populations and translocating them to 1 of 12 nearby recipient
sites (Figure S3) \citep{seddon2014}. The donor
populations included 2 of the 5 recovering populations in the frog
evolution study described above (Figure~\ref{fig-allelefrequencies} A:
population 1 and 4). Following translocation, adult survival and
recruitment of new adults were estimated from CMR surveys and counts of
tadpoles and juveniles were obtained from visual encounter surveys
(VES).

Estimates of frog survival during the 1-year period following
translocation indicate that survival was highly variable between
recipient sites, but relatively constant within recipient sites (for the
subset of sites that received multiple translocations;
Figure~\ref{fig-translocation-survival}). As such, the average survival
of frogs in the first translocation to a site was an important predictor
of survival of frogs in subsequent translocations to that site
(Figure S4). This suggests an important
effect of site characteristics and/or (site * frog) interactions on frog
survival. In addition, 1-year survival was higher for frogs translocated
later in the study period than earlier
(Table S1), which we suggest resulted
primarily from our improved ability to choose recipient sites with
higher habitat quality for \emph{R. sierrae}. Finally, Bd loads were
fairly consistent before versus after translocation and loads were
nearly always well below the level indicative of severe chytridiomycosis
and associated frog mortality (Figure S5, see
SI Appendix text for additional details)
\citep{joseph2018, vredenburg2010}.

@@

Even when incorporating (i) demographic stochasticity and (ii)
environmental stochasticity in the probability of juvenile recruitment
to the adult class \(\omega\) (the transition that we expect to be the
most subject to environmental variability in the presence of Bd),
populations with high adult survival are likely to persist over a 50
year time horizon. Our model predicted that, following a single
introduction of 40 adult individuals into a population and using the
baseline value for \(\omega\) of 0.3, the six populations with the
highest adult survival probabilities had 50-year extinction
probabilities of 0 to 0.32 (Figure~\ref{fig-viability} B). This
indicates strong potential for long-term persistence in the presence of
Bd and environmental variability in recruitment. When we decreased
\(\omega\) from 0.3 to a conservative estimate of 0.15, 50-year
extinction probability remained less than 0.5 for three populations and
was 0.79 to 0.94 for three additional populations
(Figure~\ref{fig-viability} B). In contrast, for the six populations
where yearly adult survival probability \(\sigma_{A_R} < 0.5\),
extinction probability over 50 years was always predicted to be 1. To
test the validity of our model predictions, we demonstrated that our
stochastic model could describe the general recovery trajectory of our
translocated population with the longest survey history
(Figure~\ref{fig-viability} C; population 70550, surveyed for 16 years).

@@

Disease-induced population declines are decimating global biodiversity
\citep{daszak2000}, but strategies to recover affected species are
generally lacking. Here, we tested the possibility that populations of
resistant/tolerant individuals could be used to reestablish extirpated
populations. Although previously proposed
\citep{mendelson2019, scheele2021}, this strategy has rarely been used
in the wild (see \citep{joseph2018} for a small-scale example). In
particular, although numerous studies have documented signatures of
selection in hosts following disease-induced declines
\citep[e.g.,][]{epstein2016, gignoux-wolfsohn2021}, we are unaware of
any that used individuals bearing the resulting adaptive alleles to
reestablish extirpated populations. The results of our study provide
evidence that (1) MYL frogs have experienced strong selection following
exposure to a recently-emerged lethal fungal disease and that some of
the genomic regions under selection are associated with genes related to
disease resistance/tolerance, and (2) reintroduction of individuals from
these resistant/tolerant populations allows the successful
reestablishment of viable frog populations despite ongoing disease.

Previous field studies in MYL frogs show that frog-Bd dynamics and frog
survival in the presence of Bd are fundamentally different between naive
and recovering populations. Following the arrival and establishment of
Bd in previously-naive populations, adult frogs develop high Bd loads
that lead to mass die-offs \citep{vredenburg2010}. In contrast, in
recovering populations adult frogs typically have low-to-moderate and
relatively constant Bd loads and mass die-offs are not observed
\citep[see also
Figure S5]{briggs2010, knapp2011}. The
differences in Bd load of frogs from naive and recovering populations
are also observed in controlled laboratory studies \citep[Figure
4]{knapp2016}. The observed reduced susceptibility of frogs from
recovering populations could in theory be due to several factors,
including natural selection for more resistant/tolerant genotypes and/or
acquired immunity, but until now evidence to evaluate the role of
evolution was lacking.

Results from our genomic analyses strongly suggest that natural
selection for adaptive alleles and resulting increased
resistance/tolerance is at least partially responsible for the reduced
susceptibility of frogs in recovering populations. We identified
specific alleles and genomic regions showing signatures of selection
between adjacent Bd-naive and recovering MYL frog populations,
consistent with strong selection following Bd exposure. Importantly,
some regions under selection are associated with cellular and
immunological mechanisms known to contribute to disease resistance,
including in amphibians \citep{zamudio2020}. For example, the MHC plays
an important role in immunity. In our study, we identified a region that
shows evidence of selection in recovering populations and contains eight
genes associated with either the MHC Class I or Class III regions. These
results corroborate numerous previous studies linking MHC genes to
amphibian resistance/tolerance against Bd
\citep[e.g.,][]{savage2011, bataille2015}. Similarly, the region with
the strongest indication of directional selection (as measured by
\(\pi_{diff}\)) contains the interferon-related gene GVINP1. Several
previous studies have found this gene to be differentially expressed
during Bd infection \citep[e.g.,][]{grogan2018, ellison2014} and in
populations differing in Bd susceptibility \citep{grogan2018}. This gene
is also strongly linked to disease in salmon, explaining a notable 20\%
of the resistance phenotype \citep{robledo2020, robledo2018}. We also
identified a region, characterized by high \emph{F\textsubscript{ST}}
and low \(\pi_{diff}\), that contained the complement genes C6 and C7.
This could indicate that balancing selection is acting in this region of
the genome to favor a diverse set of alleles, as is known for C6 in
humans \citep{soejima2005}. Finally, based on the analysis of individual
outlier variants, the RIN3 gene showed a consistent pattern of allele
frequency differences across all nine of the frog populations sampled in
this study, indicating consistent selection in populations distributed
across a wide geographic area. This gene is associated with inflammation
response and in \emph{Xenopus} is highly expressed in wound stumps
following amputation of tadpole tails \citep{fukazawa2009}. Also in this
analysis, the outlier variant with the lowest p-value (uncharacterized
gene LOC108802036) is located adjacent to a type I interferon gene
(Np-IFNi2) in another frog genome \citep{gan2018}, and together with
GVINP1 further indicates the importance of interferon-related genes in
this system.

@@

Despite the resistance/tolerance of adult MYL frogs in recovering
populations, individual and population-level impacts of Bd are still
evident. In an earlier study of 2 of our 12 translocated populations
\citep{joseph2018}, Bd infection and load had detectable effects on the
survival of adults and may have influenced population establishment
(sites referred to as ``Alpine'' and ``Subalpine'' in \citep{joseph2018}
are identified as ``70550'' and ``70505'' in the current study).
Applying similar analyses to all 12 of our translocated populations
would likely provide a broader perspective of the ongoing effect of Bd.
In addition to these important but relatively subtle effects of Bd on
adults, the impacts on younger life stages are more apparent. MYL frogs
immediately following metamorphosis (``metamorphs'') are highly
susceptible to Bd infection \citep{ellison2018} and as a result
experience high mortality. This high susceptibility of metamorphs is
documented for numerous species of anurans, and may result from the
poorly developed immune system characteristic of this life stage
\citep{humphries2022}. In naturally recovering and translocated MYL frog
populations, we suggest that the high mortality of metamorphs is an
important limitation on subsequent recruitment of new adults. Therefore,
although adult MYL frogs appear relatively resistant/tolerant, Bd
infection continues to have important limiting effects on recovering
populations \citep[see also][]{hollanders2022}.

The recent emergence of Bd worldwide has contributed to the decline of
hundreds of amphibian species, some of which are now extinct in the wild
\citep{scheele2019}. This unprecedented impact on global amphibian
biodiversity is compounded by the lack of any effective and broadly
applicable strategies to reverse these impacts
\citep{garner2016, knapp2022}. Importantly, in addition to the natural
recovery documented for MYL frogs \citep{knapp2016}, other species are
also showing evidence of post-epizootic recovery in the presence of Bd
\citep{scheele2017, voyles2018} and suggest the possibility of also
using animals from these recovering populations to reestablish
extirpated populations. As with MYL frogs, the feasibility and long-term
success of such efforts will depend on the availability of robust donor
populations containing individuals that have acquired via evolution the
adaptive alleles necessary to allow frog survival and population growth
in the presence of Bd. Despite the hopeful example of successful
reestablishment of MYL frogs via evolutionary rescue and
reintroductions, the challenge of recovering hundreds of Bd-impacted
amphibian species worldwide is a daunting prospect. Although we now have
a proven strategy to reestablish extirpated MYL frog populations,
recovery across their large historical range will require substantial
resources over many decades. The results of this study provide a hopeful
starting point for that endeavor and other future efforts worldwide.

Evolutionary rescue may be increasingly critical for the survival of
wildlife populations in a rapidly changing world. The importance of
evolutionary rescue is known for a wide variety of taxa
\citep[e.g.,][]{carlson2014, mills2018, oziolor2019}, including in the
specific context of disease-impacted systems \citep{searle2020}.
However, obtaining evidence in wild animal populations for all the
causal links in this process (i.e., selection for adaptive alleles
\(\rightarrow\) increase in host fitness in the presence of the stressor
\(\rightarrow\) positive population growth and recovery) is difficult in
most systems. Nonetheless, we provide a compelling example from the wild
not only that evolutionary rescue can produce individuals that harbor
adaptive alleles and allow population recovery, but importantly that
these individuals can be used in translocations to reestablish
extirpated populations. We expect that such evolutionary rescue-based
translocations will be an essential tool in wildlife conservation in an
era of accelerating global change.

@@

To identify important predictors of frog survival following
translocation, we developed linear and multilevel models in a Bayesian
framework. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}, and used default, weakly informative
priors. Starting models included all relevant population-level effects
and, when data contained possible hierarchical structure, we included
group-level effects in the full model. We evaluated model fit using
leave-one-out (``LOO'') cross-validation
\citep{gelman2013, gabry2019, vehtari2016}, as implemented by the loo
package \citep{vehtari2022}.
